Reduced diversity and increased specialization of Alaska salmon fishermen:

### 2017-01-24 salmon/paper/salmon_paper_AOS.Rmd
Ole's additions to Paper.  Sorry but I forgot to make a new branch.

---
title: 'Reduced diversity and increased specialization of Alaska salmon fishermen:
  effects on variation in individual revenues'
author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, others"
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
bibliography: CatchPortfolios_salmon.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(reshape2)
library(date)
library(ggtern)
library(mgcv)
library(fpc)
library(gridExtra)
library(glmmTMB)
library(zoo)
library(rstan)
modelname = "model-linear" # model-linear-nointeractions

@@

Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits]. Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies. [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals. While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber). Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.

Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX). For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig. A1), median revenue is increasing (across individuals, Fig. A2), and landed species diversity declining over time (averaged across individuals, Fig. A6). Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig. A4). In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy_financial_1997], and the same result may be true for salmon fishers. Further, these impacts of diversification are variable over time.

Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O. keta*], pink [*O. gorbuscha*], sockeye [*O. nerka*], coho [*O. kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears. For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas). Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively). While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species. Perhaps the best examples include fisheries that target sockeye salmon (e.g. Bristol Bay set and drift gillnet fisheries). For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].

A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals. The modern era of Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1]. Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008-1] or economic impacts [@boyce_alaska_1993], rather than to variability experienced by individual fishermen.

@@

@@

The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by [@anderson_benefits] to include time-varying effects. These time-varying effects will allow us to compare temporal trends across regions, identify periods when having diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.

>>>>

---
title: 'Reduced diversity and increased specialization of Alaska salmon fishermen:
  effects on variation in individual revenues'
author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, others"
date: ''
output:
  html_document: default
  pdf_document:
    fig_caption: yes
  word_document: default
bibliography: CatchPortfolios_salmon.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(reshape2)
library(date)
library(ggtern)
library(mgcv)
library(fpc)
library(gridExtra)
#library(glmmTMB)
library(zoo)
library(rstan)
modelname = "model-linear" # model-linear-nointeractions

@@

Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits]. Fisheries in Alaska are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies. [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals. While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber). Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.

Salmon fisheries are particularly important in Alaska both in terms of participation and value. The number of fishers participating only in salmon fisheries has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX) and overall participation, total value?, and median revenue has increased since the early 2000s (Fig. A1 and A2).  Yet the landed species diversity has declined over time (averaged across individuals, Fig. A6) and across different salmon fisheries there appears to be little correlation between diversity and revenue (Fig. A4). Thus the relationship between harvest diversity and revenue is neither simple or straight-forward. As seen in terrestrial agriculture[@purdy_financial_1997], diversification may affect the mean or CV of revenue, but not both, or the effect of diversification for revenue may vary over time [bad last sentence].

Salmon permits that allow the catch of multiple species provide an unusual opportunity for examining how the availablity of resources and fisher choice interact to affect both the total revenue and variability in revenue achieved by individual fishers. All individuals owning a single permit have the same opportunites to fish and to choose different fishing strategies. Individual fishers can therefore elect to utilize a diverse set of available species or be highly specialized on a single species. Thus the variability in fisher behavior among indivdiual fishers exist a range of strategies that allow us to understand the relationship between diversity of harvest and economic outcomes at the scale of the individual making decisions. Most previous analyses have focused on either vessel or community-level effects of diversification which are aggregate properties one step away from actual decision makers.

Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O. keta*], pink [*O. gorbuscha*], sockeye [*O. nerka*], coho [*O. kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears. For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas). Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively). Most of these gear types allow fishers the opportunity to harvest a diversity of salmon species.  In practice, though, the catch is dominated by a single species for some permits. Perhaps the best examples are the Bristol Bay set and drift gillnet fisheries where all five salmon species are present but fishers target sockeye salmon almost exclusively (REF). For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].

An additional, underappreciated consideration for understanding the relationship between diversification and fisheries value is how the availablity of different species in the environment affects the harvest diversity realized by individual fishers. For salmon one of the most imporant factors determining salmon abundance has been the expansion of artifical hatchery propagation of salmon species. Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1]. Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008-1] or economic impacts [@boyce_alaska_1993], rather than to variability experienced by individual fishermen.

@@

To understand the effects of diversification on revenue and revenue variability, we analyze single-permit salmon holders landings in Alaska between 1985 and 2015. We focus on XXX mixed species salmon fisheries in Alaska (ZZZ seine and YYY gillnet fisheries) and explore three aspects of the diversity-revenue relationship. First, we describe salmon catches - including total biomass, species composition, and diversity of catches -, fisheries participation, and revenue in Alaska over the past 30 years for each region and fishery. Second, we construct statistical time-series models to understand patterns of reveune and variability for each fishery. Specifically, we extend the methods used by [@anderson_benefits] to estimate time-varying effects of diversity on revenue and revenue variability. We then use estimated time-varying parameters to ask if the effects of diversity are constistent across time and similar among gear types and regions. Finally, we link our results to available information on economic and environmental data to uncover potential connections between the diversity-revenue relationship and longer term drivers in the region.

*Ole's proposed flow of figures.(based on Eric's figure numbers circa 1//24/2017)
  + Generally, I would add Bristol Bay sockeye at least in the time-series figures because it is such a well known and well studied fishery.  It also would be a nice contrast to the others just to show that it has effectively no diversity at all.  It may not warrant inclusion in the modeling or later analysis, but I think not mentioning the most valuable salmon fishery in Alaska is a bit odd. Also, it should be mentioned that Bristol Bay is weird and not really representative of most salmon fisheries in Alaska.
  + A Suggest putting Appendix Figs 1,2,3 as a 3 panel main text figure- pick out the focal fisheries and plot trends in participation in one panel, trends in per-capita landed weight in a second, and trends in revenue in the third (may be best to put a smoother through all of the raw data to make it palatable to the eye).  Point is that despite being in different regions and different gears, there are broad similarities in all fisheries: falling participation, rising catches, relatively flat revenue.
  + B I like figure 1, but would add Bristol Bay (see above.). May need to switch to areas as rows, and the two panels as columns.  Also I would add a total catch line to the top panel (maybe add a third panel? not sure). 
  + C Fig 2 & 3.  Suggest making the percent composition plots one array of figures and having all 6 or 7 of the diversity plots in one figure with different colors or line-types. Maybe make the diversity figure big and then the composition for each area smaller and to the side? Main point is areas differ in composition and differ in overall diversity & trend in diversity. The diversity panel may be better served in the bullet point A, showing raw trends in diversity with other raw attributes.
  + D Fig. 4. Important points here seem to be about the variation across years (some gear-regions are more variable than others) and I wonder if this would be better presented as an esimated among-year variance or SD for each area and then possibly a pairwise correlations tile-plot among areas.  Suggest putting Fig. 4 time-series in the appendix.
  + E Fig 5 and 6.  Suggest collapsing multi-panel in Fig.5 to one panel and Fig.6 to one panel. Point of Fig. 5 is choosing to add species is negative on revenue and increasingly negative in recent years.  Story of Fig 6. is a little more confused, and no clear signal of diversity on revenue variability
  + Use final figure or two to relate external information to changes in diversity-revenue or diversity-revenue variability relationship.  Possibilities: total hatchery production, price of permits.  Also could break each area into revenue-variability bi-plots and plot each fishery in that bivariate space by time block (e.g. 5 year time blocks)

Extra stuff:

@@
